Sample Length at Which Block Bootstrapping is Effective for Estimation of Parameter of Time Series

### 2023-01-18 Code/simulation.Rmd
Manuscript Change

```{r functions}
## This note checks the empirical coverage rate of CIs constructed from
## block bootstrap with series generated from an MA process

@@

do1rep <- function(n, theta, statistic, blksize, R = 1000, level = .95) {
    x <- arima.sim(list(ma = theta), n = n) / sqrt(1 + theta^2)
    bts <- boot::tsboot(x, statistic, l = blksize, sim = "fixed", R = R)
    alpha <- 1 - level
    ## returns a vector with length twice of that of mystat's returned values
    ## each pair forms an interval for one target
    pctCI <- apply(bts$t, 2, quantile, prob = c(alpha/2, 1 - alpha/2))
    empCI <- matrix(2 * bts$t0, 2, length(bts$t0), byrow = TRUE) - pctCI[c(2, 1),]
    ## BCa interval
    z0 <- qnorm(colMeans(sweep(bts$t, 2, bts$t0) < 0))
    ## thetajack <- sapply(1:n, function(i) statistic(x[-i]))
    thetajack <- sapply(1: (n / blksize),
                        function(i) statistic(x[ - ((i - 1) * blksize + 1:blksize)]))
    a <- apply(thetajack, 1, e1071::skewness) / 6
    alpha1 <- pnorm(z0 + (z0 + qnorm(alpha/2)) / (1 - a * (z0 + qnorm(alpha/2))))
    alpha2 <- pnorm(z0 + (z0 + qnorm(1 - alpha/2)) / (1 - a * (z0 + qnorm(1 - alpha/2))))
    bcaCI <- sapply(1:length(bts$t0),
                    function(i) quantile(bts$t[,i], prob = c(alpha1[i], alpha2[i])))
    ## return all intervals
    c(pctCI, empCI, bcaCI)
}

@@

```{r simulation}
nrep <- 4000
```

```{r ma = .3, n = 125}
n <- 125
blksize <- 5
theta <- .3
target <- c(0, 1, theta / (1 + theta^2))
set.seed(1234)
sim_.3_125 <- replicate(nrep, do1rep(n, theta, mystat, blksize))
mychk(sim_.3_125, target)
```

```{r ma = .3, n = 1000}
n <- 1000
blksize <- 10
theta <- .3
target <- c(0, 1, theta / (1 + theta^2))
set.seed(2345)
sim_.3_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
mychk(sim_.3_1k, target)
```

```{r ma = .3, n = 3375}
n <- 3375
blksize <- 15
theta <- .3
target <- c(0, 1, theta / (1 + theta^2))
set.seed(3456)
sim_.3_3375 <- replicate(nrep, do1rep(n, theta, mystat, blksize))
mychk(sim_.3_3375, target)
```

```{r ma = .3, n = 8000}
n <- 8000
blksize <- 20
theta <- .3
target <- c(0, 1, theta / (1 + theta^2))
set.seed(4567)
sim_.3_8k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
mychk(sim_.3_8k, target)
```

>>>>

```{r functions}
## This note checks the empirical coverage rate of CIs constructed from
## block bootstrap with series generated from an AR process

@@

do1rep <- function(n, theta, statistic, blksize, R = 1000, level = .95) {
    x <- arima.sim(list(ar = theta), n = n) * sqrt(1 - theta^2)
    bts <- boot::tsboot(x, statistic, l = blksize, sim = "fixed", R = R)
    alpha <- 1 - level

    ## Percentile interval
    mbts <- apply(bts$t, 2, mean)
    crit <- 
      apply(t(t(bts$t) - mbts), 2, quantile, prob = c(alpha/2, 1 - alpha/2)) 
    ## alpha/2 and 1 - alpha/2 critical values 
    ## of pseudo-estimate - mean of pseudo-estimates
    crit2 <- 
      apply(t(t(bts$t) - bts$t0), 2, quantile, prob = c(alpha/2, 1 - alpha/2)) 
    ## alpha/2 and 1 - alpha/2 critical values 
    ## of pseudo-estimate - original estimate

    ## interval centered around...
    estCI <- sweep(crit, 2, bts$t0, FUN = '+') # original estimate
    ## LB: bts$t0 + alpha/2 crit, UB: bts$t0 + (1 - alpha/2) crit
    mbtCI <- sweep(crit2, 2, mbts, FUN = '+') # mean of pseudo-estimates
    ## LB: mbts + alpha/2 crit2, UB: mbts + (1 - alpha/2) crit2
    altCI <- sweep(-crit[order(2:1),], 2, bts$t0, FUN = '+') # both
    ## LB: bts$t0 - (1 - alpha/2) crit, UB: bts$t0 - alpha/2 crit and
    ## LB: mbts - (1 - alpha/2) crit2, UB: mbts - alpha/2 crit2

    ## BCa interval
    z0 <- qnorm(colMeans(sweep(bts$t, 2, bts$t0) < 0))
    thetajack <- sapply(1: (n / blksize), 
                        function(i) 
                          statistic(x[ - ((i - 1) * blksize + 1:blksize)]))
    a <- apply(thetajack, 1, e1071::skewness) / 6
    alpha1 <- pnorm(
      z0 + (z0 + qnorm(alpha/2)) / (1 - a * (z0 + qnorm(alpha/2))))
    alpha2 <- pnorm(
      z0 + (z0 + qnorm(1 - alpha/2)) / (1 - a * (z0 + qnorm(1 - alpha/2))))
    crit3 <- sapply(1:length(bts$t0), function(i) 
      quantile(t(t(bts$t[,i]) - mbts[i]), prob = c(alpha1[i], alpha2[i]))) 
    ## alpha1 and alpha2 critical values 
    ## of pseudo-estimate - mean of pseudo-estimates
    crit4 <- sapply(1:length(bts$t0), function(i) 
      quantile(t(t(bts$t[,i]) - bts$t0[i]), prob = c(alpha1[i], alpha2[i])))
    ## alpha1 and alpha2 critical values 
    ## of pseudo-estimate - original estimate

    ## interval centered around...
    bcaCIest <- sweep(crit3, 2, bts$t0, FUN = '+') # original estimate
    ## LB: bts$t0 + alpha1 crit3, UB: bts$t0 + alpha2 crit3
    bcaCImbt <- sweep(crit4, 2, mbts, FUN = '+') # mean of pseudo-estimates
    ## LB: mbts + alpha1 crit4, UB: mbts + alpha2 crit4
    bcaCIalt <- sweep(-crit3[order(2:1),], 2, bts$t0, FUN = '+') # both
    ## LB: bts$t0 - alpha2 crit3, UB: bts$t0 - alpha1 crit3 and
    ## LB: mbts - alpha2 crit4, UB: mbts - alpha1 crit4

    ## return all intervals
    c(estCI, mbtCI, altCI, bcaCIest, bcaCImbt, bcaCIalt)
}

@@

```{r simulation}
nrep <- 1000
```

```{r ma = .1, n = 100}
n <- 100
blksize <- 5
theta <- .1
target <- c(0, 1, theta)
set.seed(1234)
sim_.1_1h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.1_1h,"../Data/Raw/sim_.1_1h.csv", row.names = FALSE)
mychk(sim_.1_1h, target)
```
```{r ma = .1, n = 500}
n <- 500
blksize <- 8
theta <- .1
target <- c(0, 1, theta)
set.seed(2345)
sim_.1_5h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.1_5h,"../Data/Raw/sim_.1_5h.csv", row.names = FALSE)
mychk(sim_.1_5h, target)
```
```{r ma = .1, n = 1000}
n <- 1000
blksize <- 10
theta <- .1
target <- c(0, 1, theta)
set.seed(3456)
sim_.1_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.1_1k,"../Data/Raw/sim_.1_1k.csv", row.names = FALSE)
mychk(sim_.1_1k, target)
```
```{r ma = .3, n = 100}
n <- 100
blksize <- 5
theta <- .3
target <- c(0, 1, theta)
set.seed(1234)
sim_.3_1h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.3_1h,"../Data/Raw/sim_.3_1h.csv", row.names = FALSE)
mychk(sim_.3_1h, target)
```
```{r ma = .3, n = 500}
n <- 500
blksize <- 8
theta <- .3
target <- c(0, 1, theta)
set.seed(2345)
sim_.3_5h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.3_5h,"../Data/Raw/sim_.3_5h.csv", row.names = FALSE)
mychk(sim_.3_5h, target)
```
```{r ma = .3, n = 1000}
n <- 1000
blksize <- 10
theta <- .3
target <- c(0, 1, theta)
set.seed(3456)
sim_.3_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.3_1k,"../Data/Raw/sim_.3_1k.csv", row.names = FALSE)
mychk(sim_.3_1k, target)
```
```{r ma = .5, n = 100}
n <- 100
blksize <- 5
theta <- .5
target <- c(0, 1, theta)
set.seed(1234)
sim_.5_1h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.5_1h,"../Data/Raw/sim_.5_1h.csv", row.names = FALSE)
mychk(sim_.5_1h, target)
```
```{r ma = .5, n = 500}
n <- 500
blksize <- 8
theta <- .5
target <- c(0, 1, theta)
set.seed(2345)
sim_.5_5h <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.5_5h,"../Data/Raw/sim_.5_5h.csv", row.names = FALSE)
mychk(sim_.5_5h, target)
```
```{r ma = .5, n = 1000}
n <- 1000
blksize <- 10
theta <- .5
target <- c(0, 1, theta)
set.seed(3456)
sim_.5_1k <- replicate(nrep, do1rep(n, theta, mystat, blksize))
write.csv(sim_.5_1k,"../Data/Raw/sim_.5_1k.csv", row.names = FALSE)
mychk(sim_.5_1k, target)
```

### 2023-01-18 Manuscript/manuscript.tex
Manuscript Change

Block bootstrapping is a method that can be applied to a non-stationary process to estimate a parameter. Suppose that a sample of length n of a time series is given. In regular bootstrap procedure, the new pseudo-sample of the same size as the original sample would be created by simply resampling observations from the with replacement. In the case of a time series, in order to account for the temporal dependence, the time series can be split into blocks, typically of the same size. The block should be of size l large enough to include time dependence, yet small enough to include some variance. l should also divide n. As n increases, an ideal l should also increase, but the ratio of l to n should decrease. To achieve this, l is often assigned a value as a function of n. A common function that is considered the best by much previous literature is l = n$^{1/3}$. To make things sample for this study, only perfect cubes were used for the sample length. A bootstrapped sample is created by taking a sample of n / l blocks with replacement to create a new pseudo-sample of size n. The designer of the study can choose whether the blocks overlap or not. If the blocks can never overlap, it is called a non-moving block bootstrap. In this case, the original sample is split evenly into n / l blocks, and the same number of blocks are sampled with replacement to create a bootstrapped pseudo-sample. If the blocks do overlap, it is called a moving block bootstrap. n / l samples are taken randomly from the original sample, but they do not have to be from a set of evenly spaced blocks. An estimate of the parameter (or parameters) $\hat{\theta}_{b}$ is computed from the bootstrapped sample. This procedure can be repeated many times to create a distribution of $\hat{\theta}_{b}$. In this study, 1000 bootstrapped pseudo-samples were created for each simulation. Using this distribution of $\hat{\theta}_{b}$, a 1 - $\alpha$/2 \% confidence interval for the parameter can be created. The most simplistic method for constructing this interval is the percentile confidence interval. For a parameter $\theta$, the percentile confidence interval takes the form: \[ [\hat{\theta}_{b, .025}, \hat{\theta}_{b, .975}].\] Another method is the bias-corrected and accelerated 
(BCA) confidence interval, which takes the form: \[ [\hat{\theta}_{b, \alpha_{L}}, \hat{\theta}_{b, \alpha_{U}}].\] $\alpha{L}$ and $\alpha{U}$ are the cumulative probability of $z{L}$ and $z{U}$, respectively, where:
\[z_{L} = \frac{z_{0} - z_{1 - \alpha/2}}{1 - a(z_{0} - z_{1 - \alpha/2})} + z_{0}\] and
\[z_{U} = \frac{z_{0} + z_{1 - \alpha/2}}{1 - a(z_{0} + z_{1 - \alpha/2})} + z_{0}.\] 
	Both of these intervals are good at recovering the mean and standard deviation of a temporally dependent process, but their coverage of the temporal dependence deteriorates as n increases. A potential solution to this issue is to center the intervals around either the original sample estimate $\hat{\theta}$ or the mean of all $\hat{\theta}_{b}$.

@@

\begin{figure}[tbp]
\caption{The figure below shows results for block bootstrap estimation when theta is 0. When the target was the mean and the moving method was used, the coverage rate confidence interval included .95 when the sample length was 512, indicating that even for a sample with no temporal dependence, a very large sample length of around 500 is necessary to get an accurate estimation of the mean. Similarly, when the target was the mean and the non-moving method was used, the coverage rate confidence interval included .95 when the sample length was 512, showing that a sample size of around 500 is necessary for good block bootstrap estimation. When the target was the AR(1) coefficient and the moving method was used, the parameter was over-covered - all of the coverage rate confidence interval lower bounds were greater than .95 - for all sample sizes that were observed (32, 64, 128, 256, and 512). This indicates that the variance of the parameter is being overestimated. When the target was the AR(1) coefficient and the non-moving method was used, the parameter was under-covered when the sample length was 32 and slightly over-covered when the sample length was 256. For sample lengths 64, 128, and 512, the non-moving method recovered the parameter well.}
  \centering
  \includegraphics[width=\textwidth]{theta_0}
  \caption{}
  \label{fig:theta_0}
\end{figure}

\begin{figure}[tbp]
\caption{The figure below shows results for block bootstrap estimation when theta is 0.2. When the target was the mean and the moving method was used, the coverage rate confidence interval included .95 when the sample length was 2048 and 4096, indicating that even for a sample with relatively low temporal dependence, a very large sample of around 2000 is necessary to get an accurate estimation of the mean. Similarly, when the target was the mean and the non-moving method was used, the coverage rate confidence interval included .95 when the sample length was 2048 and 4096, showing that a sample size of around 2000 is necessary for good block bootstrap estimation. When the target was the AR(1) coefficient, for both methods, the parameter was under-covered for sample lengths as large as 4096, although the moving method appeared to perform better for all sample sizes. }
  \centering
  \includegraphics[width=\textwidth]{theta_0.2}
  \caption{}
  \label{fig:theta_0.2}
\end{figure}

\begin{figure}[tbp]
\caption{The figure below shows results for block bootstrap estimation when theta is 0.4. When the target was the mean, for both methods, the coverage rate was under-covered for sample sizes as large as 8192. When the target was the AR(1) coefficient, for both methods, the parameter was under-covered for sample lengths as large as 4096, although the moving method appeared to perform better for smaller sample sizes. }
  \centering
  \includegraphics[width=\textwidth]{theta_0.4}
  \caption{}
  \label{fig:theta_0.4}
\end{figure}

>>>>

Block bootstrapping is a method that can be applied to a non-stationary process to estimate a parameter. Suppose that a sample of length n of a time series is given. In regular bootstrap procedure, the new pseudo-sample of the same size as the original sample would be created by simply resampling observations from the with replacement. In the case of a time series, in order to account for the temporal dependence, the time series can be split into blocks, typically of the same size. The block should be of size l large enough to include time dependence, yet small enough to include some variance. l should also divide n. As n increases, an ideal l should also increase, but the ratio of l to n should decrease. To achieve this, l is often assigned a value as a function of n. A common function that is considered the best by much previous literature is l = n$^{1/3}$. To make things sample for this study, only perfect cubes were used for the sample length. A bootstrapped sample is created by taking a sample of n / l blocks with replacement to create a new pseudo-sample of size n. The designer of the study can choose whether the blocks overlap or not. If the blocks can never overlap, it is called a non-moving block bootstrap. In this case, the original sample is split evenly into n / l blocks, and the same number of blocks are sampled with replacement to create a bootstrapped pseudo-sample. If the blocks do overlap, it is called a moving block bootstrap. n / l samples are taken randomly from the original sample, but they do not have to be from a set of evenly spaced blocks. An estimate of the parameter (or parameters) $\hat{\theta}_{b}$ is computed from the bootstrapped sample. This procedure can be repeated many times to create a distribution of $\hat{\theta}_{b}$. In this study, 1000 bootstrapped pseudo-samples were created for each simulation. Using this distribution of $\hat{\theta}_{b}$, a 1 - $\alpha$/2 \% confidence interval for the parameter can be created. The most simplistic method for constructing this interval is the percentile confidence interval. For a parameter $\theta$, the percentile confidence interval takes the form: \[ [\hat{\theta}_{b, .025}, \hat{\theta}_{b, .975}].\] Another method is the bias-corrected and accelerated 
(BCA) confidence interval, which takes the form: \[ [\hat{\theta}_{b, \alpha_{L}}, \hat{\theta}_{b, \alpha_{U}}].\] $\alpha{L}$ and $\alpha{U}$ are the cumulative probability of $z{L}$ and $z{U}$, respectively, where:
\[z_{L} = \frac{z_{0} - z_{1 - \alpha/2}}{1 - a(z_{0} - z_{1 - \alpha/2})} + z_{0}\] and
\[z_{U} = \frac{z_{0} + z_{1 - \alpha/2}}{1 - a(z_{0} + z_{1 - \alpha/2})} + z_{0}.\] 
	Both of these intervals are good at recovering the mean and standard deviation of a temporally dependent process, but their coverage of the temporal dependence deteriorates as n increases. A potential solution to this issue is to center the intervals around either the original sample estimate $\hat{\theta}$ or the mean of all $\hat{\theta}_{b}$. To do this, the distribution of $\delta = $\hat{\theta}_{b}$ - $\hat{\theta}$ or

@@

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.1}
  \caption{}
  \label{fig:plot_.1}
\end{figure}

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.3}
  \caption{}
  \label{fig:plot_.3}
\end{figure}

\begin{figure}[tbp]
\caption{}
  \centering
  \includegraphics[width=\textwidth]{figures/plot_.5}
  \caption{}
  \label{fig:plot_.5}
\end{figure}
